Probabilistic Policy Violation Detection with IPFWD

### 2017-02-22 paper/_main.tex
Huge clean up for final paper, paper changes

\usepackage{graphicx}
\graphicspath{ {images/} }

@@

\author{Philip Nelson and Austin Voecks% <-this % stops a space
\thanks{Supported in part by Dell EMC Isilon through grant and hardware}}

@@

\begin{abstract}
%\boldmath
When using a firewall system like IPFW to detect threats, we can end up doing a
lot of packet processing. This can negatively impact performance-sensitive
systems such as storage nodes. This paper describes a practical solution to
this problem using a load-weighted probabilistic mechanism that allows a
trade-off between perfect visibility of incoming packets and reduced impact to
system load.

@@

% TODO rewrite
\section{Introduction}

  IPFWD is a daemon for FreeBSD that updates an early rule in IPFW that has a
  chance to accept any packet. The probability of acceptance is updated over
  time and dependent on the current system load. In high performance systems
  this results in an increase in network throughput.

  This approach is based on the premise that firewall performance can be
  improved by reducing the number of rules applied to each packet. Supposing a
  white-list policy, a firewall has to apply every rule to each packet before
  it's denied.  With IPFWD, the system has a chance to accept any packet early
  and skip any further computation. Test results shows this reduces the system
  resources required to handle the same amount of traffic.

  This is a shift in mindset from typical firewalls \cite{networksecurity}.
  Instead of enforcing every part of a security policy all the time, IPFWD
  enforces the policy some of the time and extrapolates from the violations it
  encounters.  This provides insight into the violations that weren't detected.
  For this cost, you gain increased firewall performance and network throughput
  in resource bound systems. It's acceptable to allow a percentage of policy
  violations given that network traffic patterns are often repeated and the
  goal is detection and not immediate prevention. Preventative action may be
  taken later when resource requirements are lower.

  As an example, under heavy load IPFWD may immediately accept 40\% of packets.
  Supposing a port scan was initiated during this time and rules exist to block
  it, 60\% of the port scan would still be rejected and logged. Since the
  probability will fluctuate over time, IPFWD provides information in the IPFW
  logs to show the chance undetected violations occurred for each detected
  violation. IPFWD allows administrators to keep extensive rule sets that fully
  implement their security policy. Instead of having to simplify rule sets to
  increase performance, IPFWD balances policy enforcement and performance
  automatically. Under normal or light load, IPFWD will enforce the entire
  security policy 100\% of the time.

@@

    Firewalls restrict the types of incoming and outgoing network traffic
    according to a security policy. Without a security policy, administrators
    cannot know what to allow and disallow. Typical firewalls enforce the
    policy by blocking packets that do not match the rules. The default action
    for white-list firewalls is to block all traffic, additional rules describe
    exceptions. Blacklist firewalls are the opposite in that by default they
    accept all traffic and additional rules describe packet types that are not
    allowed. Since predicting and blacklisting all possible attacks is
    infeasible, white-list firewall rule sets are the accepted standard and the
    focus of this paper.

@@

    IPFW is a stateful firewall written for FreeBSD which supports both IPv4
    and IPv6. It is comprised of several components: a kernel firewall filter
    rule processor and its integrated packet accounting facility, the logging
    facility, NAT, the dummynet traffic shaper, a forward facility, a bridge
    facility, and an ipstealth facility \cite{freebsdhandbook}.

    IPFW was chosen for it's close integration with the FreeBSD operating
    system, in kernel packet filter, and general performance. Performance is
    dependent on system parameters and circumstances, but IPFW outperforms PF
    in stateful packet filtering \cite{ipfwvspf}. Stateful rules are more
    powerful and flexible than stateless rules, were no session information is
    maintained \cite{networksecurity}.  Additionally, the capabilities of IPFW
    extend beyond packet filtering into source based routing, traffic shaping
    and more.

    % First match wins rule system and probabilistic matching
    Firewall rules are treated by IPFW in a first-match-wins fashion. IPFW also
    contains built-in functionality and rule syntax for probabilistic packet
    matching. From the documentation, this feature was intended for load
    balancing and other traffic shaping tasks. It's the core of IPFWD's
    operation.

@@

  % What does it do?
  IPFWD increases firewall performance by reducing the average amount of work
  it takes to process a packet. Depending on the current system load, IPFWD
  increases or decreases the chance the IPFW will accept a given packet early.
  It does this by adding and maintaining an early rule in the IPFW rule set of
  the following form: \verb|prob 0.000 allow ip from any to any|.

  % How does it work?
  % - Probabilistic packet matching
  IPFWD's behavior is based on the assumption that the majority of network
  traffic is valid, malicious events are rare \cite{networktrafficanalysis},
  and it's sufficient to be notified of security policy violations instead of
  always stopping them. These assumptions will not hold in all environments,
  but we can leverage them for performance gains when they do.

  Since IPFW uses a first-match-wins rule system, this early rule will be
  encountered before the majority of the other rules in the rule set.  The
  probability given in this rule determines the chance that we skip processing
  all further rules, supposing that there hasn't been a match already.

  This is especially valuable for a white-list firewall rule set dealing with
  large amounts of rejected traffic; normally each rejected packet has all
  rules applied to it while IPFW looks for a match.  When no matches are
  encountered, the default rule is employed and the packet is rejected.

  In a typical data center backend network the vast majority of traffic is
  valid and going to be accepted by the firewall's rule set. However, these
  rule sets are often complex and take increasing time to match packets
  depending on the number of rules. For a rule set of length $n$, we can expect
  to a packet to encounter $\frac{n}{2}$ failed matches before matching the
  correct rule and being accepted.

@@

  % - Log extension
  While this method increases the speed the system processes packets, it
  creates a chance that packets that would have normally been rejected will get
  through the firewall. We'll refer to these types of packets as invalid for
  the purpose of discussion.

@@

        The gains provided by IPFWD are only apparent in systems were network
        throughput is matched by or out performs CPU speed. Systems that can
        already easily handle fully processing each packet will not benefit by
        reducing processing time. Given modern CPU clock speeds, systems with
        network cards slower than 1 Gb/sec will likely not benefit either.

@@

        By using IPFWD, administrators sacrifice the assurance that every
        invalid packet will be rejected.  However, they are still provided
        information as to whether some invalid packets were rejected, and that
        information can be used to infer the existence of additional invalid
        packets.

@@

  % Talk about security policy enforcement
  Security policies define what is and is not allowed in an organization . At
  the network and host level, this is implemented in part by a firewall rule
  set. A policy violation on an internal network is more serious than an
  external network. Internet facing hosts can expect to be scanned, experience
  network anomalies and be attacked by third party agents semi-regularly.

  However when this happens on an internal network, there is a strong
  possibility something else is going wrong. The usage of IPFWD is predicated
  on the assumption that it's sufficient to be notified of these events so
  further action can be taken later than to always immediately stop them.

  % What does it mean for security if you're ignoring some of the packet?
  Allowing some percentage of network packets through the firewall without
  inspection does mean that attacks that would have normally been blocked could
  succeed. In addition to performance improvements, the following factors help
  to account for this risk:

@@

      Malicious network activity and network errors are likely to be repeated
      over time \cite{beacondetection}. The longer the activity persists, the
      more likely it is to be detected by IPFW even when IPFWD is allowing a
      high percentage of the network traffic through unchecked.

@@

      Though some malicious network activity may only consist of a single
      packet, others will require multiple packets. Since IPFW works on a
      per-packet basis, malicious network activity will experience
      approximately $m\%$ packet loss where $m$ is IPFWD's probability of early
      acceptance. The exact effect this would have on malware or reconnaissance
      would depend, but generally will increase the chance they fail or produce
      invalid results.

@@

  % How do you trust this system?
  Knowledge of a network's traffic patterns and the additional log information
  provided by IPFWD can help. It allows administrators to extrapolate the
  existence of undetected network events through simple statistical models.

  % Talk about shift in mindset from "blocking attacks" to "detecting attacks"
  % - Why is this reasonable?
  % - Is this really different than traditional firewalls?
  TODO

@@

    Suppose each host in a data center's internal network is running IPFWD on
    each host, average CPU usage is $75\%$, and have rule sets blocking common
    network attacks. One of the hosts has been compromised and begins
    conducting network reconnaissance by port scanning the other hosts on the
    network.

    Firewall rules are in place to block host to host communication $m\%$ of
    the ports being scanned. This means that $1 - m\%$ of the scan would be
    allowed regardless of IPFWD's action. This percentage represents valid
    host-to-host communication as determined by the security policy.

@@

  % describe difference between reply and no reply violations
  \subsection{Security Cost}

    The chance of missing a security policy violation depends on the type of
    traffic. The following situations exist when both incoming and outgoing
    firewall rules are employed:

@@

        Here, the firewall only has one chance to catch the policy violation.
        So if the early acceptance probability is $m$, then $m\%$ of the policy
        violating packets will be allowed.

@@

        Here, the firewall has two chances to catch the policy violation.  So
        if the early acceptance probability is $m$, then $m'\%$ of the policy
        violating packets will be allowed where $m' = m * m$.

@@

  Tests were run on three platforms and five different network cards, resulting
  in five different systems. The most important parameters to our tests are
  CPU, NIC, and ENV. Descriptions of each system are given in the tables
  below:\\

@@

  Virtualisation makes it more difficult to have the same confidence in the
  results as with the physical machines. Tests on the Virtualised system were
  repeated additional times to account for the variability of the underlying
  physical hardware and contention between virtual machines.

  The WWU and Virtualised tests were conducted on stock FreeBSD 11 without
  additional network performance configuration applied. Isilon OneFS does
  contain additional network performance tuning and demonstrates the additional
  gains IPFWD can provide with careful FreeBSD 11 kernel configuration.

@@

    % Explain tools
    Multiple network performance tools were explored but the final results were
    compiled using Netperf. Netperf was chosen because it was readily available
    on all test platforms, provides built-in CPU utilization measurement, and a
    wide range of test types \cite{netperf}.

    % Explain tests
    The test type UDP\_STREAM was chosen over TCP\_STREAM for simplicity and to
    preserve CPU cycles for IPFW. Raw network throughput and CPU performance is
    the target metric, which can be difficult to assess with TCP
    \cite{netperf}.

    Netperf is single threaded, which allows easier comparison of results
    between systems with different numbers of cores. Additionally, all tests
    were run while the system was idle to minimize contention over system
    resources.

@@

    CPU utilization measurement is the most difficult parameter to measure in
    these tests. There are many factors that can account for variability
    between tests and jitter in the results. The Netperf manual describes these
    challenges and the variety of techniques used on different systems.
    \cite{netperf}

    Netperf was run in single threaded mode, which explains why throughput
    wasn't closer to the maximum throughput for the NIC on each system. The
    FreeBSD network stack allows multithreaded packet processing by utilizing
    multiple queues \cite{freebsdhandbook}. Since we're indirectly testing the
    performance of IPFW through Netperf, there are two different CPU
    measurements to account for: kernel time and user time. Since IPFW runs in
    the kernel, it's CPU time is classified as kernel time.

    Netperf was able to measure CPU utilization on the Virtualized and WWU
    systems, but didn't work on the Isilon systems. Instead, \textit{ps} was
    used. These tests are comparable since both measured user CPU utilization,
    not kernel CPU utilization.

    In systems where the throughput of the NIC exceeds what the CPU can supply,
    CPU utilization is high and throughput is lower than the NIC maximum (eg:
    Virtualised system). Conversely, when CPU availablility exceeds NIC
    throughput, throughput is high and CPU utilization is low (eg: WWU 1G
    system).

@@

    These graphs show the interaction between CPU utilization and network
    throughput for each system with respect to IPFWD's early acceptance
    probability. The dashed lines are CPU utilization, solid is network
    throughput in Mb/second. Higher acceptance probability means a greater
    chance that a given packet is immediately accepted without further
    processing.

    The upward trend in throughput on the x-axis shows that as the early
    acceptance probability increases, so does throughput. The CPU usage shows
    that this throughput is gained without additional CPU utilization, meaning
    the system is spending fewer cycles to process each packet on average.

    The lack of any trend in the WWU 1G graph is expected and addressed in the
    \textit{Discussion} section.

    \begin{center}
      \includegraphics[width=0.45\textwidth]{cpu_isilon40}
      \includegraphics[width=0.45\textwidth]{cpu_isilon10}
      \includegraphics[width=0.45\textwidth]{cpu_wwu1}
      \includegraphics[width=0.45\textwidth]{cpu_virtual10}
    \end{center}

@@

    These graphs show network the ratio of throughput to CPU utilization for
    each system where $ratio(x) = throughput(x) / cpu(x)$. The data being
    presented here is the same as the previous graphs, shown differently. The
    upward trend over the x-axis shows that as the probability of early
    acceptance increases, the system is spending less CPU time to process
    packets on average.

    The dashed line shows the line of best fit for the throughput to CPU
    utilization ratio and shows more clearly the upward trend.

    \begin{center}
      \includegraphics[width=0.45\textwidth]{toc_isilon40}
      \includegraphics[width=0.45\textwidth]{toc_isilon10}
      \includegraphics[width=0.45\textwidth]{toc_wwu1}
      \includegraphics[width=0.45\textwidth]{toc_virtual10}
    \end{center}

@@

    In CPU bound systems, IPFWD reduces the number of cycles it takes to
    process each packet, allowing more packets to be processed in the same time
    and increasing throughput. This is shown most clearly in the Virtualised
    system tests, which have the weakest CPU but 10Gb Ethernet NICs. Regardless
    of test type, the receiving system is always at 100\% CPU utilization.
    However, as the probability to accept increases in IPFWD the throughput
    also increases.  This clearly shows that IPFWD decreases the CPU load
    required to process incoming traffic.

@@

  % Packet sampling
  Other papers \cite{exploitingpacketsampling, analysisnetflow,
  monitoringpacketsampling} on packet sampling have been focused on detecting
  and classifying threats, not performance enhancement. However, their
  approaches to packet sampling and the implications addressed provided the
  idea for IPFWD.

  % IPFW performance
  The BSD community has produced several in depth investigations into network
  and firewall performance \cite{ipfwvspf,optimizingfreebsd}. Their work
  informed the tests run to measure IPFWD's performance benefits.

@@

  % Firewall rule reordering so we can "fail as early as possible"
  Ideally, a white-list firewall's rule set would be constructed so that the
  most commonly applied matches are reached earlier rather than later. IPFWD
  accomplishes this task by artificial early stopping. Firewall rule reordering
  is a could be another solution to this problem. Given logs of common network
  traffic for the system, statistical analysis could show which rules are
  matched more often than others. With this information, the rule set could be
  reordered so that the most commonly applied rules are earliest.

  Other work \cite{redundancydetection} has been done in removing redundant
  rules to increase firewall performance. It's possible that a combination of
  redundancy checking and priority reordering could provide further performance
  gains.

  % creating baseline invalid packet % so we can extrapolate even when we don't
  % see any invalid packets 
  Machine learning could be applied to packet samples to more accurately
  determine appropriate bounds on the early acceptance probability. Different
  systems and circumstances may require more or less stringent packet checking.
  Environments with very high confidence in network security might raise the
  lower bound of the acceptance rate to improve performance. Likewise, less
  confident environments may lower the upper bound to ensure an acceptable
  percentage of packets are always checked.

  % have detection of bad traffic reduce acceptance rate
  Future versions of IPFWD could benefit from additional modifiers to the early
  acceptance probability. For instance, the system could reduce the early
  acceptance probability whenever packets are dropped by the firewall under the
  assumption that invalid traffic likely precedes more invalid traffic.

@@

  In performance sensitive computing environments, complex firewall rules
  combined with high network throughput can result in heavy CPU load. IPFWD
  addresses this problem by providing a graceful trade off between packet
  visibility, total security policy enforcement and system performance.

@@

\appendices
\section{Additional Information}
  The source code for IPFWD, additional test results, and development cycle
  information may be found at \texttt{https://???}.

% use section* for acknowledgement
%\section*{Acknowledgment}
% TODO

>>>>

\usepackage{caption}
\usepackage{graphicx}
\graphicspath{ {images/} }

@@

\author{Austin Voecks and Philip Nelson% <-this % stops a space
\thanks{Supported in part by Dell EMC Isilon through grant and hardware}}

@@

\begin{abstract}
%\boldmath
When using a firewall system like IPFW to detect threats, we can end up doing a
lot of packet processing. This can negatively impact performance-sensitive
systems such as storage nodes in data centers. This paper describes a
practical solution to this problem using a load-weighted probabilistic
mechanism that allows a trade-off between perfect visibility of incoming
packets and reduced impact to system load.

@@

\section{Introduction}

  IPFWD is a daemon for FreeBSD that updates an early rule in the IP Firewall
  (IPFW) that has a chance to accept any packet. The probability of early
  acceptance is updated over time and dependent on the current system load. In
  high performance systems this can result in an increase in network
  throughput.

  This approach is based on the premise that firewall performance can be
  improved by reducing the number of rules applied to each packet. Supposing a
  white-list policy, a firewall must apply every rule to a packet before
  it's denied.  With IPFWD, the system has a chance to accept any packet early
  and skip any further computation. Test results show this reduces the 
  resources required to handle the same amount of traffic in some systems.

  This is a shift in mindset from typical firewalls \cite{networksecurity}.
  Instead of enforcing every part of the security policy all the time, IPFWD
  enforces the policy some of the time provides additional information so the
  missed violations can be inferred. For this cost, you gain increased firewall
  performance and network throughput in resource bound systems. IPFWD works
  under the assumption that it's acceptable to allow a percentage of policy
  violations given that network traffic patterns are often repeated and the
  goal is detection, not immediate prevention. Administrative action may be
  taken later when resource requirements are lower.

  As an example, under heavy load IPFWD may immediately accept 40\% of packets.
  Some of those packets may have been malicious. Supposing a port scan was
  initiated during this time and rules exist to block it, at least 60\% of the
  port scan would still be rejected and logged. Since the early acceptance
  probability will fluctuate over time, IPFWD provides information in the IPFW
  logs to show the chance undetected violations occurred for each detected
  violation. IPFWD allows administrators to keep extensive rule sets that fully
  implement their security policy. Instead of having to simplify rule sets to
  increase performance, IPFWD balances policy enforcement and performance
  automatically. Under normal or light load, IPFWD will enforce the entire
  security policy 100\% of the time.

@@

    Firewalls restrict the types of incoming and outgoing network traffic
    according to a security policy. Without a security policy, administrators
    cannot know what to allow and disallow when constructing firewall rules.
    Typical firewalls enforce the policy by blocking packets that do not match
    any rules. The default action for white-list firewalls is to block all
    traffic, additional rules describe exceptions. Blacklist firewalls are the
    opposite, by default they accept all traffic and additional rules describe
    packet types that are not allowed. Since predicting and blacklisting all
    possible attacks is not possible in practice, white-list firewall rule sets
    are the accepted standard and the focus of this paper.

@@

    IPFW is a stateful firewall written for FreeBSD which supports both IPv4
    and IPv6. It is comprised of several components: a kernel firewall filter
    rule processor and its integrated packet accounting facility, a logging
    facility, NAT, the dummynet traffic shaper, a forward facility, a bridge
    facility, and an ipstealth facility \cite{freebsdhandbook}.

    IPFW was chosen for its close integration with the FreeBSD operating
    system, kernel packet filter, and performance. Performance is dependent on
    system parameters and circumstances, but IPFW outperforms PF in stateful
    packet filtering \cite{ipfwvspf}. Stateful rules are more powerful and
    flexible than stateless rules, were no session information is maintained
    \cite{networksecurity}.  Additionally, the capabilities of IPFW extend
    beyond packet filtering into source based routing, traffic shaping and
    more. It's for these reasons that IFPW was chosen over PF.

    % First match wins rule system and probabilistic matching
    Firewall rules are treated by IPFW in a first-match-wins fashion. IPFW also
    contains built-in functionality and rule syntax for probabilistic packet
    matching. From the documentation, this feature was intended for load
    balancing and other traffic shaping tasks. It's the core of IPFWDs
    operation.

@@

  % What does it do?
  IPFWD increases firewall performance by reducing the average amount of work
  it takes to process a packet. Depending on the current system load, IPFWD
  increases or decreases the chance the IPFW will accept a given packet early.
  It does this by adding and maintaining an early rule in the IPFW rule set of
  the following form: 
  \begin{center}
      \verb|prob 0.000 allow ip from any to any|
  \end{center}
  which allows any IP packet, incoming or outgoing with a chance equal to the probability given.

  % How does it work?
  % - Probabilistic packet matching
  IPFWDs behavior is based on the assumption that the majority of network
  traffic is valid, malicious events are rare \cite{networktrafficanalysis},
  and it's sufficient to be notified of security policy violations instead of
  always stopping them. These assumptions will not hold in all environments,
  but we can leverage them for performance gains when they do.

  Since IPFW uses a first-match-wins rule system, this early rule will be
  encountered before the majority of the other rules in the rule set.  The
  probability given in this rule determines the chance that IPFWD skips
  processing all further rules, supposing that there has not been a match
  already.

  This is especially valuable for a white-list firewall rule set dealing with
  large amounts of rejected traffic; normally each rejected packet has all
  rules applied to it in order to find a match. When no matches are
  encountered, the default rule is employed and the packet is rejected.

  In a typical data center backend network, the vast majority of traffic is
  valid and going to be accepted by the firewalls rule set. However, these
  rule sets are often complex and take increasing time to match packets
  depending on the number of rules. For a rule set of length $n$, we can expect
  a packet to encounter on average $\frac{n}{2}$ failed matches before matching
  the correct rule and being accepted.

@@

  % - Log extension
  While this method increases the speed the system processes packets, it
  creates a chance that packets that would have normally been rejected will get
  through the firewall.

@@

        The gains provided by IPFWD are only apparent in systems were network
        throughput is matched by or out performs CPU speed. Systems that can
        already easily handle fully processing each packet will not benefit by
        reducing processing time. Given modern CPU clock speeds, systems with
        network cards slower than 1 Gb/sec will likely not benefit from IPFWD.

@@

        By using IPFWD, administrators sacrifice the assurance that every
        invalid packet will be rejected.  However, they are still provided
        information as to whether some invalid packets were rejected, and that
        information can be used to infer the existence of additional invalid
        packets. Environments that cannot allow this relaxation in security
        policy enforcement will no benefit from IPFWD.

@@

  % Talk about security policy enforcement
  Security policies define what is and is not allowed in an organization. At
  the network and host level, this is implemented in part by a firewall rule
  set. A policy violation on an internal network is more serious than an
  external network. Internet facing hosts can expect to be scanned, experience
  network anomalies, and be attacked by third party agents more often than
  internal hosts.

  However, when these events do occur on an internal network, there is a strong
  possibility something else is wrong. Unexpected network activity on a fully
  controlled internal network indicates compromise or misconfiguration, both of
  which are serious problems. The usage of IPFWD is predicated on the
  assumption that it's sufficient to be notified of these events so further
  action can be taken later.

  % What does it mean for security if you're ignoring some of the packet?
  Allowing some percentage of network packets through the firewall without
  inspection does mean that attacks that would have normally been blocked could
  succeed. The following factors help to account for this risk:

@@

      Malicious network activity such as malware beaconing and network
      reconnaissance, as well as general network errors are likely to be
      repeated over time \cite{beacondetection}. The longer the activity
      persists, the more likely it is to be detected by IPFW, even when IPFWD
      is allowing a high percentage of the network traffic through unchecked.

@@

      Though some malicious network activity may only consist of a single
      packet, others will require multiple packets. Since IPFW works on a
      per-packet basis, malicious network activity will experience
      approximately $(1 - m)\%$ packet loss where $m$ is IPFWDs probability of early
      acceptance. The exact effect this would have on malware or reconnaissance
      would depend, but generally will increase the chance of failure or
      invalid results.

@@

  % How do you trust this system?

  % Talk about shift in mindset from "blocking attacks" to "detecting attacks"
  % - Why is this reasonable?
  % - Is this really different than traditional firewalls?

@@

    Suppose each host in a data centers internal network is running IPFW and
    IPFWD on each host, average CPU usage is $75\%$, and have rule sets
    blocking common network attacks. One of the hosts has been compromised and
    begins conducting network reconnaissance by port scanning the other hosts
    on the network.

    Firewall rules are in place to block host-to-host communication to $m\%$ of
    the ports being scanned. This means that $1 - m\%$ of the scan would be
    allowed regardless of IPFWDs action. This percentage represents valid
    host-to-host communication as determined by the security policy.

@@

  % describe difference between reply and no reply violations
  Thus, the chance of missing a security policy violation depends on the type
  of traffic. The following situations exist when both incoming and outgoing
  firewall rules are employed:

@@

      Here, the firewall only has one chance to catch the policy violation.
      So, if the early acceptance probability is $m$, then $m\%$ of the policy
      violating packets will be allowed.

@@

      Here, the firewall has two chances to catch the policy violation. So,
      if the early acceptance probability is $m$, then $m'\%$ of the policy
      violating packets will be allowed where $m' = m * m$.

@@

  Tests were run on three platforms and four different network cards, referred
  to as four different systems. The most important parameters to these tests
  are CPU, NIC, and ENV. Descriptions of each system are given in the tables
  below:\\

@@

  Virtualisation made it more difficult to have the same confidence in the
  results as with the physical machines. Tests on the Virtualised system were
  repeated additional times to account for the variability of the underlying
  physical hardware and contention between virtual machines.

  The WWU and Virtualised tests were conducted on stock FreeBSD 11 without
  additional network performance configuration applied. The Isilon systems
  contained additional network performance tuning and demonstrates the
  additional gains IPFWD can provide with careful FreeBSD 11 kernel
  configuration.

@@

    % Explain tools
    Multiple network performance tools were explored but the final results were
    compiled using Netperf. Netperf was chosen because it was readily available
    on all test platforms, provides built-in CPU utilization measurement, has a
    wide range of test types \cite{netperf}, and is used elsewhere in the BSD
    community for testing.

    % Explain tests
    The test type UDP\_STREAM was chosen over TCP\_STREAM for simplicity and to
    preserve CPU cycles for IPFW. Raw network throughput and CPU performance
    was the target metric, which can be difficult to assess with TCP
    \cite{netperf}.

    Netperf is single threaded, which allowed easier comparison of results
    between systems with different numbers of cores. Additionally, all tests
    were run while the system was idle to minimize contention over system
    resources.

@@

    CPU utilization measurement was the most difficult parameter to measure in
    these tests. There are many factors that can account for variability
    between tests and jitter in the results. The Netperf manual describes these
    challenges and the variety of techniques used on different systems to
    measure CPU usage \cite{netperf}.

    Netperf was run in single threaded mode, which explains why throughput was
    not closer to the maximum throughput for the NIC on each system. The
    FreeBSD network stack allows multithreaded packet processing by utilizing
    multiple queues meaning the kernel can spread the work done by IPFW over
    multiple cores \cite{freebsdhandbook}. Since we were indirectly testing the
    performance of IPFW through Netperf, there were two different CPU
    measurements to account for: kernel time and user time. Since IPFW runs in
    the kernel, its CPU time is classified as kernel time.

    Netperf was able to measure CPU utilization on the Virtualised and WWU
    systems, but didn't work on the Isilon systems. Instead, \textit{ps} was
    used. These tests are comparable since both measured user CPU utilization,
    not kernel CPU utilization. Thus, IPFWs CPU usage was measured indirectly
    through the CPU usage of netperf.

    In systems where the throughput of the NIC exceeds what the CPU can supply,
    CPU utilization is high and throughput is lower than the NIC maximum (eg:
    the Virtualised system). Conversely, when CPU availability exceeds NIC
    throughput, throughput is high and CPU utilization is low (eg: the WWU 1G
    system).

@@

    The graphs in figure \ref{fig:isiloncpu40}, \ref{fig:isiloncpu10}, \ref{fig:wwucpu1}, and \ref{fig:virtualcpu10} show the interaction between CPU utilization and network
    throughput for each system with respect to IPFWDs early acceptance
    probability. The dashed lines are CPU utilization; solid is network
    throughput in Mb/second. A higher acceptance probability means a greater
    chance that a given packet was immediately accepted without further
    processing.

    The upward trend in throughput on the x-axis shows that as the early
    acceptance probability increases, so does throughput. The CPU usage shows
    that this additional throughput was gained without additional CPU
    utilization, meaning the system was spending fewer cycles to process each
    packet on average.

    The lack of any trend in the WWU 1G graph was expected and addressed in the
    \textit{Discussion} section.

    \begin{figure}[h]
      \includegraphics[width=0.45\textwidth]{cpu_isilon40}
      \caption{Throughput \& CPU for Isilon 40G system}
      \label{fig:isiloncpu40}
    \end{figure}
    \begin{figure}[h]
      \includegraphics[width=0.45\textwidth]{cpu_isilon10}
      \caption{Throughput \& CPU for Isilon 10G system}
      \label{fig:isiloncpu10}
    \end{figure}
    \begin{figure}[h]
      \includegraphics[width=0.45\textwidth]{cpu_wwu1}
      \caption{Throughput \& CPU for WWU 1G system}
      \label{fig:wwucpu1}
    \end{figure}
    \begin{figure}[h]
      \includegraphics[width=0.45\textwidth]{cpu_virtual10}
      \caption{Throughput \& CPU for Virtualised 10G system}
      \label{fig:virtualcpu10}
    \end{figure}

    %\begin{center}
    %  \includegraphics[width=0.45\textwidth]{cpu_isilon40}
    %  \includegraphics[width=0.45\textwidth]{cpu_isilon10}
    %  \includegraphics[width=0.45\textwidth]{cpu_wwu1}
    %  \includegraphics[width=0.45\textwidth]{cpu_virtual10}
    %\end{center}

@@

    The graphs in figures \ref{fig:isilontoc40}, \ref{fig:isilontoc10}, \ref{fig:wwutoc1}, and \ref{fig:virtualtoc10}  show network the ratio of throughput to CPU utilization for
    each system where $ratio(x) = throughput(x) / cpu(x)$. The data being
    presented here is the same as the previous graphs, merely shown
    differently. The upward trend over the x-axis shows that as the probability
    of early acceptance increases, the system is spending less CPU time to
    process packets on average.

    The dashed line shows the line of best fit for the throughput to CPU
    utilization ratio and shows the upward trend more clearly than the raw
    data.

    \begin{figure}[h]
      \includegraphics[width=0.45\textwidth]{toc_isilon40}
      \caption{Throughput/CPU for Isilon 40G system}
      \label{fig:isilontoc40}
    \end{figure}
    \begin{figure}[h]
      \includegraphics[width=0.45\textwidth]{toc_isilon10}
      \caption{Throughput/CPU for Isilon 10G system}
      \label{fig:isilontoc10}
    \end{figure}
    \begin{figure}[h]
      \includegraphics[width=0.45\textwidth]{toc_wwu1}
      \caption{Throughput/CPU for WWU 1G system}
      \label{fig:wwutoc1}
    \end{figure}
    \begin{figure}[h]
      \includegraphics[width=0.45\textwidth]{toc_virtual10}
      \caption{Throughput/CPU for Virtualised 10G system}
      \label{fig:virtualtoc10}
    \end{figure}

    %\begin{center}
    %  \includegraphics[width=0.45\textwidth]{toc_isilon40}
    %  \includegraphics[width=0.45\textwidth]{toc_isilon10}
    %  \includegraphics[width=0.45\textwidth]{toc_wwu1}
    %  \includegraphics[width=0.45\textwidth]{toc_virtual10}
    %\end{center}

@@

    In CPU bound systems, IPFWD reduces the number of cycles it takes to
    process each packet, allowing more packets to be processed in the same time
    and increasing throughput. This is shown most clearly in the Virtualised
    system tests, which have the weakest CPU but 10Gb Ethernet NICs. Regardless
    of test type, the receiving system is always at 100\% CPU utilization.
    However, as the probability to accept increases in IPFWD the throughput
    also increases.  This clearly shows that IPFWD decreases the CPU time
    required to process incoming traffic.

@@

  % Packet sampling
  Other papers \cite{exploitingpacketsampling, analysisnetflow,
  monitoringpacketsampling} on packet sampling have been focused on detecting
  and classifying threats, not performance enhancement. However, their
  approaches to packet sampling and the implications addressed provided the
  inspiration for IPFWD.

  % IPFW performance
  The BSD community has produced several in depth investigations into network
  and firewall performance \cite{ipfwvspf,optimizingfreebsd}. Their work
  informed the tests run here.

@@

  % Firewall rule reordering so we can "fail as early as possible"
  Ideally, a white-list firewalls rule set would be constructed so that the
  most commonly applied matches are reached sooner rather than later. IPFWD
  accomplishes this task by artificial early stopping. Firewall rule reordering
  could be another solution to this problem. Given logs of common network
  traffic for the system, statistical analysis could show which rules are
  matched more often than others. With this information, the rule set could be
  reordered so that the most commonly applied rules are earliest.

  Other work \cite{redundancydetection} has been done in removing redundant
  rules to increase firewall performance. It is possible that a combination of
  redundancy checking and priority reordering could provide further performance
  gains.

  % creating baseline invalid packet % so we can extrapolate even when we don't
  % see any invalid packets 
  Machine learning could be applied to packet samples to more accurately
  determine appropriate bounds on the early acceptance probability. Different
  systems and environments may require more or less stringent packet checking.
  Environments with very high confidence in network security might raise the
  lower bound of the acceptance rate to improve performance. Likewise, less
  confident environments may lower the upper bound to ensure an acceptable
  percentage of packets are always checked.

  % have detection of bad traffic reduce acceptance rate
  Future versions of IPFWD could benefit from additional modifiers to the early
  acceptance probability. For instance, the system could reduce the early
  acceptance probability whenever packets are dropped by the firewall, under
  the assumption that invalid traffic likely precedes more invalid traffic.

@@

  In performance sensitive computing environments, complex firewall rules
  combined with high network throughput can result in heavy CPU load, limiting
  system performance. IPFWD addresses this problem by providing a graceful
  trade off between packet visibility, total security policy enforcement and
  system performance.

@@

\appendices
\section{Additional Information}
  The source code for IPFWD, additional test results, and development cycle
  information may be found at \texttt{https://TODO}.
